---
title: "Tidy Tuesday for July 30th, 2025"
subtitle: "Netflix Engagement"
author: "Irene Morse"
#date: "2025-07-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup and Introduction

```{r, warning=FALSE, message=FALSE}
library("ggplot2")
library("lubridate")
library("tidyverse")
#library("jtools")
#library("kableExtra")
```

```{r, message=FALSE}
movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/movies.csv')
shows <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/shows.csv')
```

[This week's Tidy Tuesday dataset](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-07-15/readme.md) deals with the topic of funding for the British Library, which is the national library of the UK. The British Library has faced significant fluctuations in funding over the years, and I would like to investigate some possible causes for these fluctuations.

```{r}
movies$type <- "movie"
shows$type <- "show"
netflix_df <- rbind(movies, shows)
head(netflix_df)
```


## Step 1: Visualize Funding Changes Over Time

```{r}
summary(netflix_df$release_date)
```


```{r}
ggplot(netflix_df, aes(x=type, y=views)) + 
  geom_boxplot(outlier.colour="red")+theme_minimal()
```


```{r}
# exclude recently released media
print(nrow(netflix_df))
netflix_df <- subset(netflix_df, release_date <= as.Date("2024-06-30"))
print(nrow(netflix_df))
```


```{r}
netflix_df$release_season <- NA
row <- 1
for(month in lubridate::month(netflix_df$release_date)){
  if(month >= 6 & month <= 8){
    netflix_df$release_season[row] <- "summer"
  }else if(month >= 9 & month <= 11){
    netflix_df$release_season[row] <- "fall"
  }else if(month == 12 | month <= 2){
    netflix_df$release_season[row] <- "winter"
  }else if(month >= 3 & month <= 5){
    netflix_df$release_season[row] <- "spring"
  }else{
    print("Some error has occurred")
    netflix_df$release_season[row] <- "error"
  }
  row = row + 1
}

summary(as.factor(netflix_df$release_season))
```

```{r}
ggplot(netflix_df, aes(x=release_date, y=views, group=type, color=release_season))+geom_point()+scale_colour_manual(values=c("darkorange","green4","gold","darkblue"))+theme_minimal()
```


```{r}
# simulate data for number of netflix members
# assumption: growth in views reflects growth in members
# it is not driven by existing members watching more
# calculate daily growth rate
netflix_df <- netflix_df %>% arrange(release_date) %>% group_by(release_date) %>% mutate(avg_views_today=mean(views))
all_dates <- unique(netflix_df$release_date)
daily_growth <- data.frame(release_date=all_dates, daily_growth=NA)
for(i in 1:(length(all_dates)-1)){
  og <- which(netflix_df$release_date==all_dates[i])
  new <- which(netflix_df$release_date==all_dates[i+1])
  og_views <- unique(netflix_df$avg_views_today[og])
  new_views <- unique(netflix_df$avg_views_today[new])
  daily_growth$daily_growth[i] <- 100*(new_views - og_views)/og_views
}
print(daily_growth)
```


```{r}
ggplot(daily_growth, aes(x=release_date, y=daily_growth))+geom_line()
```

```{r}
netflix_df$release_month <- format(netflix_df$release_date, "%Y-%m")
netflix_df <- netflix_df %>% arrange(release_month) %>% group_by(release_month) %>% mutate(avg_views_month=mean(views))
all_months <- unique(netflix_df$release_month)
monthly_growth <- data.frame(release_month=all_months, monthly_growth=NA)
for(i in 1:(length(all_months)-1)){
  og <- which(netflix_df$release_month==all_months[i])
  new <- which(netflix_df$release_month==all_months[i+1])
  og_views <- unique(netflix_df$avg_views_month[og])
  new_views <- unique(netflix_df$avg_views_month[new])
  monthly_growth$monthly_growth[i] <- 100*(new_views - og_views)/og_views
}
print(monthly_growth)
```

```{r}
ggplot(monthly_growth, aes(x=ymd(paste(release_month, "1", sep="-")), y=monthly_growth))+geom_line()
```





```{r}
netflix_df$release_year <- format(netflix_df$release_date, "%Y")
netflix_df <- netflix_df %>% arrange(release_year) %>% group_by(release_year) %>% mutate(avg_views_year=mean(views))
all_years <- unique(netflix_df$release_year)
yearly_growth <- data.frame(release_year=all_years, yearly_growth=NA)
for(i in 1:(length(all_years)-1)){
  og <- which(netflix_df$release_year==all_years[i])
  new <- which(netflix_df$release_year==all_years[i+1])
  og_views <- unique(netflix_df$avg_views_year[og])
  new_views <- unique(netflix_df$avg_views_year[new])
  yearly_growth$yearly_growth[i] <- 100*(new_views - og_views)/og_views
}
print(yearly_growth)
```


```{r}
ggplot(yearly_growth, aes(x=as.numeric(release_year), y=yearly_growth))+geom_line()
```


```{r}
ggplot(netflix_df, aes(x=release_season, y=views, color=type))+geom_point()+geom_jitter(width=0.2)+scale_colour_manual(values = c("red","darkgray"))+theme_minimal()
```





> **Hypothesis 1**: Movies and shows released in the summer will receive more views. (mechanism = more free time)



## Step 2: Add Info about Political Parties and Visualize



## Step 4: Test Hypotheses

```{r}
# create lagged econ indicators
bl_funding$lagged_gdp_trillions <- c(bl_funding$gdp_trillions[2:nrow(bl_funding)], NA)
bl_funding$lagged_unemployment <- c(bl_funding$unemployment[2:nrow(bl_funding)], NA)
head(bl_funding[,c("year","gdp_trillions","lagged_gdp_trillions","unemployment","lagged_unemployment")])
```

Before running my regression, I have created lagged versions of the GDP variable and unemployment variable. This is to account for the fact that funding changes likely occur *after* changes in economic health. There is typically a lag as governments adjust to new information about the economy. For example, the government in 2007 must rely on economic indicators from 2006. Therefore lagged variables will allow me to more accurately model the impact of the economy on funding for the library.

```{r}
# linear model
funding_model <- lm(nominal_gbp_millions ~ lagged_gdp_trillions + lagged_unemployment + pm_party, data=bl_funding)
jtools::summ(funding_model)
```

The linear regression indicates that all three variables of interest have a statistically significant relationship with funding for the British Library. Interestingly, the health of the economy appears to be inversely related with British Library funding; higher GDP and lower unemployment are associated with *decreased* funding for the British Library. It is not immediately clear to me why this would be the case, but perhaps during stronger economic periods Parliament is able to justify decreases in government funding by arguing that the library can rely on other sources of funding. The political party variable is more readily interpretable with the Labour party clearly associated with increased funding for the library.

## Conclusion

My proposed hypotheses proved fruitful for understanding changes in British Library funding over time. It is clear that political party priorities do affect funding, with the Labour party more willing to fund the library. Additionally, the health of the economy has a notable impact, though it is the opposite of what I might have guessed. Further investigation of more micro-level economic indicators or of Parliamentary debate regarding British Library funding might help reveal the mechanisms behind this inverse relationship.